---
title: "Game of Transposons: The Bells"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#loading in packages
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(data.table)
library(formattable)
```

```{r,echo=FALSE}
# Defining variables to make it more readable later on

rpom <- 'Ruegeria_pomeroyi'
rpomColor <- 'cornflowerblue'
mari <- 'Marivivens'
mariColor <- 'orange'
vib <- 'Vib_70'
vibColor <- 'forestgreen'
diatom <- 'T_pseudo'
diatomColor <- 'khaki4'
bacteria <- 'Bacteria'
bacteriaColor <- 'gray80'
compCommColor <- 'gray25'
totalCFUColor <- 'black'
```


```{r, echo=FALSE, fig.align='center', fig.height=6, fig.width=6}
# Transform data from CFU to CFU/ml
# from 1/2 YTSS and 1/2 YTSS + KAN25
coloniesMl <- function(rawCFUdata){
    rawCFUdata <- rawCFUdata %>% 
    mutate(cfu.ml = CFU*Dilution/Volume) %>% 
    mutate(cfuKAN.ml = CFU_KAN*Dilution/Volume) %>% 
    ungroup()
}

# Calculate mean and standard error of the mean of CFU/ml 
# and CFU_KAN/ml separate and then combine
CFUmean <- function(CFUml){
    cfuMean <- CFUml %>%
        select(Condition, Replicate, Day, Organism, cfu.ml) %>%
        na.omit() %>%
        group_by(Condition, Day, Organism, Replicate) %>%
        summarize(CFU.ml = mean(cfu.ml)) %>% 
        group_by(Condition, Day, Organism) %>% 
        summarize(mean.CFU.ml=round(mean(CFU.ml)), sem.CFU.ml=sd(round(CFU.ml))/sqrt(length(CFU.ml)))%>%
        ungroup() 
    cfuKANmean <- CFUml %>%
        select(Condition, Replicate, Day, Organism, cfuKAN.ml) %>%
        na.omit() %>%
        group_by(Condition, Day, Organism, Replicate) %>%
        summarize(CFUKAN.ml = mean(cfuKAN.ml)) %>% 
        group_by(Condition, Day, Organism) %>% 
        summarize(mean.cfuKAN.ml=round(mean(CFUKAN.ml)), sem.cfuKAN.ml=sd(round(CFUKAN.ml))/sqrt(length(CFUKAN.ml)))%>%
        ungroup()
    full_join(cfuMean, cfuKANmean, by=c('Condition', 'Organism', 'Day'))
}


# Calculate mean Cells/ml from flow data taken on CyAN 
meanCellsml <- function(rawFlowdata){
    flowMean <- rawFlowdata %>%
        select(Condition, Replicate, Day, Organism, Cells.ml) %>%
        na.omit() %>%
        group_by(Condition, Day, Organism) %>%
        summarize(mean.Cells.ml=mean(round(Cells.ml)), sem.Cells.ml=sd(round(Cells.ml))
                  /sqrt(length(Cells.ml)))%>%
        ungroup()
}

# Read in CFU and flow data, transforming with above functions
# Need to add in Day 0 for condition 4/6 since these
# R pom colonies were inhibited by Vibrio, thus
# Taking the mean of all other time zeros and placing # it here

rawCFUcounts <- read.table(file='Data/GOTn_CFUcounts_v2.csv', header=TRUE, sep=',')
CFUml <- coloniesMl(rawCFUcounts)
meanCFUml <- CFUmean(CFUml)
meanCFUml <- meanCFUml[order(meanCFUml$Day), ]

flowCounts <- read.table(file='Data/GOTn_flowCounts_v2.csv', header=TRUE, sep=',')
meanFlow <- meanCellsml(flowCounts)
```

```{r, echo=FALSE}
# Fixing the problem that there were no Rpom counts on day zero of conditions 4 & 6
# We just set the CFUs here equal to the mean of the inoculum from all other conditions


condition4RpomT0 <- meanCFUml$Day==0 & meanCFUml$Organism==rpom & meanCFUml$Condition=='Vibrio'
condition6RpomT0 <- meanCFUml$Day==0 & meanCFUml$Organism==rpom & meanCFUml$Condition=='CompleteCommunity'
allRpomT0 <- na.omit(CFUml$cfu.ml[CFUml$Day=='0' & CFUml$Organism==rpom])

meanCFUml$mean.CFU.ml[condition4RpomT0] <- mean(allRpomT0)
meanCFUml$sem.CFU.ml[condition4RpomT0] <- sd(allRpomT0)/sqrt(length(allRpomT0))

meanCFUml$mean.CFU.ml[condition6RpomT0] <- mean(allRpomT0)
meanCFUml$sem.CFU.ml[condition6RpomT0] <- sd(allRpomT0)/sqrt(length(allRpomT0))
```


```{r, echo=FALSE}
#Total CFUs from means then propagate error

totalCFUS <- meanCFUml %>% 
  group_by(Condition, Day) %>% 
  summarize(total.meanCFU= sum(mean.CFU.ml), total.sem=sqrt(sum(sem.CFU.ml^2, na.rm=TRUE)))

```


```{r, echo=FALSE}
attach(meanCFUml)

meanCFU <- mean.CFU.ml[Organism==rpom]
meanCFUT0 <- mean.CFU.ml[Organism==rpom & Day==0]
meanCFUT4 <- mean.CFU.ml[Organism==rpom & Day==4]
meanCFUT8 <- mean.CFU.ml[Organism==rpom & Day==8]

semCFUT0 <- sem.CFU.ml[Organism==rpom & Day==0]
semCFUT4 <- sem.CFU.ml[Organism==rpom & Day==4]
semCFUT8 <- sem.CFU.ml[Organism==rpom & Day==8]


rpomAloneCFU <- mean.CFU.ml[Condition=='RpomAlone' & Organism==rpom]
semRpomAloneCFU <- sem.CFU.ml[Condition=='RpomAlone' & Organism==rpom]
rpomAloneTimeCFU <- Day[Condition=='RpomAlone' & Organism==rpom]

vibRpomCFU <- mean.CFU.ml[Condition=='Vibrio' & Organism==rpom]
semVibRpomCFU <- sem.CFU.ml[Condition=='Vibrio' & Organism==rpom]
vibCFU <- mean.CFU.ml[Condition=='Vibrio' & Organism==vib]
semVibCFU <- sem.CFU.ml[Condition=='Vibrio' & Organism==vib]
vibTimeCFU <- Day[Condition=='Vibrio' & Organism==vib]
vibRpomTimeCFU <- Day[Condition=='Vibrio' & Organism==rpom]

mariRpomCFU <- mean.CFU.ml[Condition=='Marivivens' & Organism==rpom]
semMariRpomCFU <- sem.CFU.ml[Condition=='Marivivens' & Organism==rpom]
mariCFU <- mean.CFU.ml[Condition=='Marivivens' & Organism==mari]
semMariCFU <- sem.CFU.ml[Condition=='Marivivens' & Organism==mari]
mariTimeCFU <- Day[Condition=='Marivivens' & Organism==mari]
mariRpomTimeCFU <- Day[Condition=='Marivivens' & Organism==rpom]

compCommRpomCFU <- mean.CFU.ml[Condition=='CompleteCommunity' & Organism==rpom]
semCompCommRpomCFU <- sem.CFU.ml[Condition=='CompleteCommunity' & Organism==rpom]
compCommMariCFU <- mean.CFU.ml[Condition=='CompleteCommunity' & Organism==mari]
semCompCommMariCFU <- sem.CFU.ml[Condition=='CompleteCommunity' & Organism==mari]
compCommVibCFU <- mean.CFU.ml[Condition=='CompleteCommunity' & Organism==vib]
semCompCommVibCFU <- sem.CFU.ml[Condition=='CompleteCommunity' & Organism==vib]
compCommVibTimeCFU <- Day[Condition=='CompleteCommunity' & Organism==vib]
compCommMariTimeCFU <- Day[Condition=='CompleteCommunity' & Organism==mari]
compCommRpomTimeCFU <- Day[Condition=='CompleteCommunity' & Organism==rpom]

detach(meanCFUml, unload=TRUE)

```


Total bacteria and diatoms were enumerated using flow cytometry (cells/ml; dotted line samples), while individual bacterial taxa were enumerated by CFU. Standard error of the mean is displayed for each treatment as black bars (n=4 for flow cytometry)


```{r,echo=FALSE}
# Reading in functions for growth curves that I like to make
niceGrowthCurve <- function(time, populationDensity, yLim, yLab=expression('Cells or CFU ml' ^-1),  timeTicks, timeLabs, popTicks, popLabs){
  xLab <- 'Days'
  axesCol <- 'gray25'
  plot(time, populationDensity, las=1, xlab=xLab, ylab=yLab, log='y',  type='n', frame.plot=FALSE,  yaxt='n', xaxt='n', ylim=yLim, col.lab=axesCol)
  axis(1, at=timeTicks, labels=timeLabs, col=axesCol, col.axis=axesCol, cex.axis=1.3, cex.lab=1.3)
  axis(2, at=popTicks, labels=popLabs, las=1, col=axesCol, col.axis=axesCol, cex.axis=1.3, cex.lab=1.3)
}


oneMemberGrowthCurve <- function(timePop1, pop1, pop1SEM, pop1Col, lineType=1, point=16){
  points(timePop1, pop1, col=pop1Col, pch=point, cex=1.5)
  lines(timePop1, pop1, col=pop1Col, lty=lineType, lwd=1.5)
  segments(x0=timePop1, y0=pop1+pop1SEM, y1=pop1-pop1SEM, col=pop1Col)
}

```


```{r, echo=FALSE}
#Defining the flow cytometry objects as the treatmentOrganism

attach(meanFlow)

tnBacteriaFlow <-  mean.Cells.ml[Condition=='RpomAlone' & Organism==bacteria]
semTnBacteriaFlow <- sem.Cells.ml[Condition=='RpomAlone' & Organism==bacteria]
tnDiatomFlow <- mean.Cells.ml[Condition=='RpomAlone' & Organism==diatom]
semTnDiatomFlow <- sem.Cells.ml[Condition=='RpomAlone' & Organism==diatom]
tnBacteriaTimeFlow <- Day[Condition=='RpomAlone' & Organism==bacteria]
tnDiatomTimeFlow <- Day[Condition=='RpomAlone' & Organism==diatom]

vibBacteriaFlow <-  mean.Cells.ml[Condition=='Vibrio' & Organism==bacteria]
semVibBacteriaFlow <- sem.Cells.ml[Condition=='Vibrio' & Organism==bacteria]
vibDiatomFlow <- mean.Cells.ml[Condition=='Vibrio' & Organism==diatom]
semVibDiatomFlow <- sem.Cells.ml[Condition=='Vibrio' & Organism==diatom]
vibBacteriaTimeFlow <- Day[Condition=='Vibrio' & Organism==bacteria]
vibDiatomTimeFlow <- Day[Condition=='Vibrio' & Organism==diatom]

mariBacteriaFlow <-  mean.Cells.ml[Condition=='Marivivens' & Organism==bacteria]
semMariBacteriaFlow <- sem.Cells.ml[Condition=='Marivivens' & Organism==bacteria]
mariDiatomFlow <- mean.Cells.ml[Condition=='Marivivens' & Organism==diatom]
semMariDiatomFlow <- sem.Cells.ml[Condition=='Marivivens' & Organism==diatom]
mariBacteriaTimeFlow <- Day[Condition=='Marivivens' & Organism==bacteria]
mariDiatomTimeFlow <- Day[Condition=='Marivivens' & Organism==diatom]

allBacteriaFlow <-  mean.Cells.ml[Condition=='CompleteCommunity' & Organism==bacteria]
semAllBacteriaFlow <- sem.Cells.ml[Condition=='CompleteCommunity' & Organism==bacteria]
allDiatomFlow <- mean.Cells.ml[Condition=='CompleteCommunity' & Organism==diatom]
semAllDiatomFlow <- sem.Cells.ml[Condition=='CompleteCommunity' & Organism==diatom]
allBacteriaTimeFlow <- Day[Condition=='CompleteCommunity' & Organism==bacteria]
allDiatomTimeFlow <- Day[Condition=='CompleteCommunity' & Organism==diatom]
detach(meanFlow, unload=TRUE)
```


##Growth Curve
```{r, echo=FALSE, fig.align='center', fig.height=4, fig.width=12}
xAxis <- seq(0,8, by=2)
xAxisLong <- seq(0,16, by=2)
xAxisLabs <- c('0','','4','','8')
xAxisLongLabs <- c('0','','4','','8','','12','','16')

ylim <- c(2e2,2e7)
yAxis <-  c(1e3, 1e4, 1e5, 1e6, 1e7)
yAxisLab <- c(expression('10'^'3'), '', expression('10'^'5'), '',expression('10'^'7'))

#pdf(file='growthcurve_20220316.pdf', height=3, width=9, useDingbats=FALSE)

par(mfrow=c(1,4), mar=c(5,4,4,1))

#Rpom Alone
niceGrowthCurve(xAxis, yAxis, ylim, timeTicks=xAxis, timeLabs=xAxisLabs, popTicks=yAxis, popLabs=yAxisLab)


oneMemberGrowthCurve(tnBacteriaTimeFlow, tnBacteriaFlow, semTnBacteriaFlow, bacteriaColor)
oneMemberGrowthCurve(tnDiatomTimeFlow, tnDiatomFlow, semTnDiatomFlow, diatomColor)
oneMemberGrowthCurve(rpomAloneTimeCFU, rpomAloneCFU, semRpomAloneCFU, rpomColor,lineType=0, point=17)




#Vibrio

niceGrowthCurve(xAxis, yAxis, ylim, yLab='', timeTicks=xAxis, timeLabs=xAxisLabs, popTicks=yAxis, popLabs=yAxisLab)

oneMemberGrowthCurve(vibBacteriaTimeFlow, vibBacteriaFlow, semVibBacteriaFlow, bacteriaColor)
oneMemberGrowthCurve(vibDiatomTimeFlow, vibDiatomFlow, semVibDiatomFlow, diatomColor)


oneMemberGrowthCurve(totalCFUS$Day[totalCFUS$Condition=='Vibrio'], totalCFUS$total.meanCFU[totalCFUS$Condition=='Vibrio'], totalCFUS$total.sem[totalCFUS$Condition=='Vibrio'], totalCFUColor, lineType=0, point=15)

oneMemberGrowthCurve(vibRpomTimeCFU, vibRpomCFU, semVibRpomCFU, rpomColor, lineType=0, point=17)
oneMemberGrowthCurve(vibTimeCFU, vibCFU, semVibCFU, vibColor, lineType=0, point=17)


legend(1.2, 0.3e4, legend=c('Total Bacteria' , 'T. pseudonana'), col=c(bacteriaColor, diatomColor), cex=0.8, pch=16, lty=1, bty='n', title='Flow Cytometry', seg.len=3)


#Marivivens

niceGrowthCurve(xAxis, yAxis, ylim, yLab='', timeTicks=xAxis, timeLabs=xAxisLabs, popTicks=yAxis, popLabs=yAxisLab)

oneMemberGrowthCurve(mariBacteriaTimeFlow, mariBacteriaFlow, semMariBacteriaFlow, bacteriaColor)
oneMemberGrowthCurve(mariDiatomTimeFlow, mariDiatomFlow, semMariDiatomFlow, diatomColor)

oneMemberGrowthCurve(totalCFUS$Day[totalCFUS$Condition=='Marivivens'], totalCFUS$total.meanCFU[totalCFUS$Condition=='Marivivens'], totalCFUS$total.sem[totalCFUS$Condition=='Marivivens'], totalCFUColor, lineType=0, point=15)

oneMemberGrowthCurve(mariRpomTimeCFU, mariRpomCFU, semMariRpomCFU, rpomColor, lineType=0, point=17)
oneMemberGrowthCurve(mariTimeCFU, mariCFU, semMariCFU, mariColor, lineType=0, point=17)


legend(1.2, 0.6e4, legend=c('R. pomeroyi DSS3' , 'V. hepatarius HF70', 'M. donghaensis HF1', 'Total Bacteria'), col=c(rpomColor, vibColor, mariColor, totalCFUColor), pch=c(17,17,17,15), cex=0.8, lty=0, bty='n', title='CFU', seg.len=3)

#CompleteCommunity

niceGrowthCurve(xAxis, yAxis, ylim, yLab='', timeTicks=xAxis, timeLabs=xAxisLabs, popTicks=yAxis, popLabs=yAxisLab)

oneMemberGrowthCurve(allBacteriaTimeFlow, allBacteriaFlow, semAllBacteriaFlow, bacteriaColor)
oneMemberGrowthCurve(allDiatomTimeFlow, allDiatomFlow, semAllDiatomFlow, diatomColor)


oneMemberGrowthCurve(totalCFUS$Day[totalCFUS$Condition=='CompleteCommunity'], totalCFUS$total.meanCFU[totalCFUS$Condition=='CompleteCommunity'], totalCFUS$total.sem[totalCFUS$Condition=='CompleteCommunity'], totalCFUColor, lineType=0, point=15)

oneMemberGrowthCurve(compCommRpomTimeCFU, compCommRpomCFU, semCompCommRpomCFU, rpomColor, lineType=0, point=17)
oneMemberGrowthCurve(compCommRpomTimeCFU, compCommVibCFU, semCompCommVibCFU, vibColor, lineType=0, point=17)
oneMemberGrowthCurve(compCommMariTimeCFU, compCommMariCFU, semCompCommMariCFU, mariColor, lineType=0, point=17)



#dev.off()
```

### Expansion Factors

```{r, echo=FALSE}
# expansion factors for Rpom as the only bacteria makes it easy

expansionCalc <- function(initialFC, finalFC, semInitialFC, semFinalFC){
  
  x <- initialFC
  y <- finalFC
  
  expansion <- y/x
  
  dx_x <- semInitialFC/x
  dy_y <- semFinalFC/y
  
  uncertainty <- expansion*sqrt((dx_x)^2*(dy_y)^2)
  
  expansion_uncertainty <- c(vector(),c(expansion, uncertainty))
}


rpomAloneExp <- expansionCalc(tnBacteriaFlow[1], tnBacteriaFlow[5], semTnBacteriaFlow[1], semTnBacteriaFlow[5])[1]
semRpomAloneExp <- expansionCalc(tnBacteriaFlow[1], tnBacteriaFlow[5], semTnBacteriaFlow[1], semTnBacteriaFlow[5])[2]

rpomAloneExpCFU <- expansionCalc(rpomAloneCFU[1], rpomAloneCFU[3], semRpomAloneCFU[1], semRpomAloneCFU[3])[1]
semRpomAloneExpCFU <- expansionCalc(rpomAloneCFU[1], rpomAloneCFU[3], semRpomAloneCFU[1], semRpomAloneCFU[3])[2]


# to get expansion factors for conditions with multiple bacteria
# Rpom CFUs / Total Bacteria CFUs * Total Bacteria FC = Rpom FC
# expansion factor is then Rpom FC final / Rpom FC initial

percentMultiBact <- function(rPom, bact2, bact3){
  x <- rPom
  y <- bact2
  z <- bact3
  
  total <- x/(x+y+z)
}

expansionCalc_noUncert <- function(initialFC, finalFC){
  
  x <- initialFC
  y <- finalFC
  
  expansion <- y/x
}

  
rpomVibPercent <- percentMultiBact(vibRpomCFU, vibCFU, bact3=0)
rpomMariPercent <- percentMultiBact(mariRpomCFU, mariCFU, bact3=0)
rpomAllPercent <- percentMultiBact(compCommRpomCFU, compCommVibCFU, compCommMariCFU)




#these are all of specifically rpom in each treatment then multiplied by the flowcytometry data (0d and 8d) to get cells/ml of rpom from the percentage of rpom cells based on cfus

vibExp <-  expansionCalc_noUncert(vibBacteriaFlow[1]*rpomVibPercent[1], vibBacteriaFlow[5]*rpomVibPercent[3])


mariExp <-  expansionCalc_noUncert(mariBacteriaFlow[1]*rpomMariPercent[1], mariBacteriaFlow[5]*rpomMariPercent[3])


allExp <- expansionCalc_noUncert(allBacteriaFlow[1]*rpomAllPercent[1], allBacteriaFlow[5]*rpomAllPercent[3])

print(c('RpomAlone Expansion=', round(rpomAloneExp, 2),'Vibrio Expansion=', round(vibExp,2) , 'Marivivens Expansion=', round(mariExp,2), 'CompleteCommunity Expansion', round(allExp, 2)))
```
 
## Number of generations over 8 days


```{r, echo=FALSE}



#Calculating the # of generations from the FC data, for rpom Alone just use the flow data, for rpom in the other communities, use the %CFU rpom * FC data

generations <- function(FC, percentRpom){
  (log10(FC[5]*percentRpom[3]) - log10(FC[1]*percentRpom[1]))/0.301

}


generationTable <- matrix(nrow=4, ncol=2)
generationTable[1, 1] <- 'rpomAlone Generations'
generationTable[2, 1] <- 'Vibrio rpom Generations'
generationTable[3, 1] <- 'Marivivens rpom Generations'
generationTable[4, 1] <- 'Complete Comm rpom Generations'
generationTable[1,2] <- round((log10(tnBacteriaFlow[5]) - log10(tnBacteriaFlow[1]))/0.301,2)
generationTable[2,2] <- round(generations(vibBacteriaFlow, rpomVibPercent),2)
generationTable[3,2] <- round(generations(mariBacteriaFlow, rpomMariPercent), 2)
generationTable[4,2] <- round(generations(allBacteriaFlow, rpomAllPercent), 2)

print(generationTable)

paste('Mean Generations=', mean(as.numeric(generationTable[1:4, 2])))
```

## Number of generations over 4 days
```{r, echo=FALSE}
# Calculating the number of generations from time 0 to 4 for just Rpom CFUs
#propagation of standard error of the mean
errorProp <- function(sem, mean){
  sqrt((0.434*(sem[2]/mean[2]))^2 + (0.434*(sem[1]/mean[1]))^2)
}

gen4Table <- matrix(nrow=4, ncol=3)
gen4Table[1, 1] <- 'rpomAlone Generations'
gen4Table[2, 1] <- 'Vibrio rpom Generations'
gen4Table[3, 1] <- 'Marivivens rpom Generations'
gen4Table[4, 1] <- 'Complete Comm rpom Generations'
gen4Table[1, 2] <- round((log10(rpomAloneCFU[2])-log10(rpomAloneCFU[1]))/0.301, 3)
gen4Table[2, 2] <- round((log10(vibRpomCFU[2])-log10(vibRpomCFU[1]))/0.301, 3)
gen4Table[3, 2] <- round((log10(mariRpomCFU[2])-log10(mariRpomCFU[1]))/0.301, 3)
gen4Table[4, 2] <- round((log10(compCommRpomCFU[2])-log10(compCommRpomCFU[1]))/0.301, 3)

gen4Table[1,3] <- round(errorProp(semRpomAloneCFU, rpomAloneCFU), 3)
gen4Table[2,3] <- round(errorProp(semVibRpomCFU, vibRpomCFU), 3)
gen4Table[3,3] <- round(errorProp(semMariRpomCFU, mariRpomCFU), 3)
gen4Table[4,3] <- round(errorProp(semCompCommRpomCFU, compCommRpomCFU), 3)

errorProp(semRpomAloneCFU, rpomAloneCFU)

print(gen4Table)
paste('Mean Generations=', mean(as.numeric(gen4Table[1:4, 2])))


```

# Fitness analysis

## Overview of fitness profiles
```{r, warning=FALSE, echo=FALSE, fig.align='center', fig.height=4, fig.width=9}
meanFitness <- fread('meanFitness_20220601.csv', header=TRUE)
significantFitness <- fread('significant_genes_grouped_Metabolism_filtered_20220601.csv', header=TRUE)
table1 <- fread('Table1_20220601.csv', header=TRUE) %>% 
    inner_join(., significantFitness, by='spo')

#Select only significant genes to later be counted as fit increases or decreases
vibSig <- significantFitness[Vib_sig == 'yes']
vibSigTable <- table1[Vib_sig=='yes']
mariSig <- significantFitness[Mari_sig == 'yes']
mariSigTable <- table1[Mari_sig=='yes']
commSig <- significantFitness[CC_sig == 'yes']
commSigTable <- table1[CC_sig=='yes']

legendColors <- c(rpomColor, vibColor, mariColor, compCommColor)

fitTicks <- c(0, 0.5, 1, 1.5, 2)
fitLabs <- c('0', '','1','', '2')

overallFitPlot <- function(meanFitX=meanFitness$mean_rpomAlone, meanFitY=meanFitness$mean_vib, SigFitY_X, SigFitY_Y, SigTableX, SigTableY, xLab, yLab, ycol){
    x <- cor.test(meanFitX, meanFitY, method='spearman')
    segmentCol <- 'grey45'
    axesCol <- 'gray25'
    plot(meanFitX, meanFitY, pch=16, col='grey90', xlab=xLab, ylab=yLab, las=1, bty='n', xaxt='n', yaxt='n',  cex=0.6, xlim=c(0,2), ylim=c(0,2), col.lab=axesCol)
    
    points(SigFitY_X, SigFitY_Y, pch=16, col=alpha(ycol,0.3), cex=0.6)
    points(SigTableX, SigTableY, pch=16,col=ycol, cex=0.6)
    rug(SigFitY_X, side=1, pos=c(2,0), ticksize=0.1, lwd=0.15, col=rpomColor)
    rug(SigFitY_Y, side=2, col=ycol, pos=c(2,0), ticksize=0.1, lwd=0.15)
    axis(1, at=fitTicks, labels=fitLabs, col=axesCol, col.axis=axesCol)
    axis(2, at=fitTicks, labels=fitLabs, col=axesCol, col.axis=axesCol)
    segments(x0=0, y0=0, x1=2, y1=2, col= segmentCol, lty=2, lwd=0.5)
    segments(x0=-1, y0=1, x1=1, y1=1, col=segmentCol, lty=2, lwd=0.5)
    text(0.5, 1.9, paste('rho=',round(x$estimate, 2)), cex=0.6)
}

print(c('rpomAlone against vibrio spearman:', cor.test(meanFitness$mean_rpomAlone, meanFitness$mean_vib, method='spearman')[3:4]))
print(c('rpomAlone against marivivens spearman:', cor.test(meanFitness$mean_rpomAlone, meanFitness$mean_mari, method='spearman')[3:4]))
print(c('rpomAlone against Complete Commuinity spearman:', cor.test(meanFitness$mean_rpomAlone, meanFitness$mean_CC, method='spearman')[3:4]))


#pdf('fitnessOverview20220601.pdf', height=3, width=9,useDingbats=FALSE)
par(mfrow=c(1,4), mar=c(5,4,4,1))

overallFitPlot(meanFitness$mean_rpomAlone, meanFitness$mean_vib, vibSig$mean_rpomAlone, vibSig$mean_vib, vibSigTable$mean_rpomAlone, vibSigTable$mean_vib, 'W(Rp)','W(Rp+V)', vibColor)


overallFitPlot(meanFitness$mean_rpomAlone, meanFitness$mean_mari, mariSig$mean_rpomAlone, mariSig$mean_mari,  mariSigTable$mean_rpomAlone, mariSigTable$mean_mari, 'W(Rp)','W(Rp+M)', mariColor)

overallFitPlot(meanFitness$mean_rpomAlone, meanFitness$mean_CC, commSig$mean_rpomAlone, commSig$mean_CC,commSigTable$mean_rpomAlone, commSigTable$mean_CC, 'W(Rp)','W(Rp+V+M)', compCommColor)
#dev.off()
```

## Binomial test to compare proportion of increased vs decreased mutants

```{r, echo=FALSE}
vibSigIncrease <- length(vibSig$Vib_fold_diff[vibSig$Vib_fold_diff > 0 & !is.na(vibSig$Vib_fold_diff)])
vibSigDecrease <- length(vibSig$Vib_fold_diff[vibSig$Vib_fold_diff < 0 & !is.na(vibSig$Vib_fold_diff)])
vibTotal <- length(vibSig$Vib_fold_diff)


mariSigIncrease <- length(mariSig$Mari_fold_diff[mariSig$Mari_fold_diff > 0 & !is.na(mariSig$Mari_fold_diff)])
mariSigDecrease <- length(mariSig$Mari_fold_diff[mariSig$Mari_fold_diff < 0 & !is.na(mariSig$Mari_fold_diff)])
mariTotal <- length(mariSig$Mari_fold_diff)

commSigIncrease <- length(commSig$CC_fold_diff[commSig$CC_fold_diff > 0 & !is.na(commSig$CC_fold_diff)])
commSigDecrease <- length(commSig$CC_fold_diff[commSig$CC_fold_diff < 0 & !is.na(commSig$CC_fold_diff)])
commTotal <- length(commSig$CC_fold_diff)

prop.test(c(vibSigIncrease, vibSigDecrease), c(vibTotal, vibTotal))
prop.test(c(mariSigIncrease, mariSigDecrease), c(mariTotal, mariTotal))
prop.test(c(commSigIncrease, commSigDecrease), c(commTotal, commTotal))
```




# Testing the fitness increases and decreases

```{r, echo=FALSE, fig.align='center'}


#pdf('log2FC_Whist_20220601.pdf', width=5, height=5)
hist(vibSig$Vib_fold_diff, seq(-6.5, 6.6, length.out = 125), col=vibColor, xlim=c(-6.5, 6.5), ylim=c(0,15), border='white', xlab='log2_FC Rp+V', main='')
hist(mariSig$Mari_fold_diff, seq(-6.5, 6.6, length.out = 125), col=mariColor, xlim=c(-6.5, 6.5), ylim=c(0,15), border='white', xlab='log2_FC Rp+M', main='')
hist(commSig$CC_fold_diff, seq(-6.6, 6.6, length.out = 125), col=compCommColor, xlim=c(-6.5, 6.5), ylim=c(0,15), border='white', xlab='log2_FC Rp+V+M', main='')
#dev.off()

# wilcox.test(vib_inc_dW, cc_inc_dW, conf.int = TRUE)
# wilcox.test(vib_inc_dW, mari_inc_dW, conf.int = TRUE)
# wilcox.test(mari_inc_dW, cc_inc_dW, conf.int = TRUE)
# median(vib_inc_dW)
# median(mari_inc_dW)
# median(cc_inc_dW)
# 
# wilcox.test(vib_dec_dW, cc_dec_dW, conf.int = TRUE)
# wilcox.test(vib_dec_dW, mari_dec_dW, conf.int = TRUE)
# wilcox.test(mari_dec_dW, cc_dec_dW, conf.int = TRUE)
# median(vib_dec_dW)
# median(mari_dec_dW)
# median(cc_dec_dW)
# 
# 
# 
# diffTab <- data.table(Treatment='Rp+V', dW=vibSig$Vib_diff) %>% 
#   rbind(data.table(Treatment='Rp+M', dW=mariSig$Mari_diff)) %>% 
#   rbind(data.table(Treatment='Rp+V+M', dW=commSig$CC_diff))
# 
# #test of -dW
# kruskal.test(diffTab$dW[diffTab$dW<0] ~ diffTab$Treatment[diffTab$dW<0])
# 
# #test of +dW
# kruskal.test(diffTab$dW[diffTab$dW>0] ~ diffTab$Treatment[diffTab$dW>0])
```





# Ammonium/Nitrate growth curves

```{r, warning=FALSE, echo=FALSE, fig.align='center', fig.height=5, fig.width=5}
plateOD <- fread('Data/NitrogenUsage_GrowthCurves_CBS.csv') %>%  melt(., id=('Time'), variable.name='Well', value.name='OD600')


plateMap <- fread('Data/NitrogenUsage_GrowthCurves_CBS_platelayout.csv')

growthData <- inner_join(plateOD, plateMap, by='Well')
meanCells <- function(rawData) {
    rawData <- rawData %>% 
        na.omit() %>% 
        group_by(Time, Organism, Condition) %>% 
        summarise(mean_OD=mean(OD600), sem_OD = sd(OD600)/sqrt(length(OD600)))
}

x <- meanCells(growthData)
x$mean_OD[x$Organism!='Blank'] <- x$mean_OD[x$Organism!='Blank'] - x$mean_OD[x$Organism=='Blank']

# Blank transformed data
blankedGrowthCurve <- x

attach(blankedGrowthCurve)

# Define each strain/fluorescent organism used for this plate
rpomNH <- Organism=='Rpom' & Condition=='NH'
rpomNO <- Organism=='Rpom' & Condition=='NO'
rpomNoN <- Organism=='Rpom' & Condition=='NoN'
vibNH <- Organism=='Vib' & Condition=='NH'
vibNO <- Organism=='Vib' & Condition=='NO'
vibNoN <- Organism=='Vib' & Condition=='NoN'
mariNH <- Organism=='Mari' & Condition=='NH'
mariNO <- Organism=='Mari' & Condition=='NO'
mariNoN <- Organism=='Mari' & Condition=='NoN'


OD <- function(strain, color) {
    points(Time[strain], mean_OD[strain], pch=16, cex=0.8, col=color)
    lines(Time[strain], mean_OD[strain], col=color, lwd=1.2)
    segments(Time[strain], y0=mean_OD[strain]-sem_OD[strain], y1=mean_OD[strain]+sem_OD[strain], col=color, lwd=1.2)
}


#pdf('Nitrate_Ammonium_GC_20211209.pdf', height=8, width=10, useDingbats = FALSE)

plot(Time[rpomNH], mean_OD[rpomNH], type='n',  las=1, xlab='Time(Hr)', ylab='OD600', main='Growth on Ammonium', bty='n', ylim=c(0, 0.4),cex=1.5, col='gray80')
OD(rpomNH, rpomColor)
OD(vibNH, vibColor)
OD(mariNH, mariColor)


plot(Time[rpomNO], mean_OD[rpomNO], type='n',  las=1, xlab='Time(Hr)', ylab='OD600', main='Growth on Nitrate', bty='n', ylim=c(0,0.4), cex=1.5, col='gray80')
OD(rpomNO, rpomColor)
OD(vibNO, vibColor)
OD(mariNO, mariColor)


plot(Time[rpomNoN], mean_OD[rpomNoN], type='n',  las=1, xlab='Time(Hr)', ylab='OD600', main='Growth on Glucose no Nitrogen', bty='n', ylim=c(0,0.4), cex=1.5, col='gray80')
OD(rpomNoN, rpomColor)
OD(vibNoN, vibColor)
OD(mariNoN, mariColor)
legend(30, 0.3, legend=c('R. pomeroyi WT', 'Vibrio', 'Marivivens'), col=c(rpomColor, vibColor, mariColor), bty='n', pch=16)
#dev.off()
detach(blankedGrowthCurve)
```

# Growth on malate, spermidine, glycolate

```{r}

glycolate_OD <- fread('Data/Glycolate_Malate_GrowthCurves_CBS.csv') %>%  melt(., id=('Time'), variable.name='Well', value.name='OD600')
glycolate_OD$Well <- as.integer(glycolate_OD$Well)
glycolate_layout <- fread('Data/Glycolate_Malate_GrowthCurves_CBS_platelayout.csv')



glycolate_growthData <- inner_join(glycolate_OD, glycolate_layout , by='Well')


glycolate_mean <- meanCells(glycolate_growthData)

# Blank transformed data

glycolate_mean$mean_OD[glycolate_mean$Organism!='Blank' & glycolate_mean$Condition=='Glycolate'] <- glycolate_mean$mean_OD[glycolate_mean$Organism!='Blank' & glycolate_mean$Condition=='Glycolate'] - glycolate_mean$mean_OD[glycolate_mean$Organism=='Blank' & glycolate_mean$Condition=='Glycolate']


attach(glycolate_mean)
rpomGlycolate <- Organism=='Rpom' & Condition=='Glycolate'

vibGlycolate <- Organism=='Vib' & Condition=='Glycolate'

mariGlycolate <- Organism=='Mari' & Condition=='Glycolate'




OD <- function(strain, color) {
    points(Time[strain], mean_OD[strain], pch=16, cex=0.8, col=color)
    lines(Time[strain], mean_OD[strain], col=color, lwd=1.2)
    segments(Time[strain], y0=mean_OD[strain]-sem_OD[strain], y1=mean_OD[strain]+sem_OD[strain], col=color, lwd=1.2)
}


#pdf('glycolateMalate_growth.pdf')
plot(Time[rpomGlycolate], mean_OD[rpomGlycolate], type='n',  las=1, xlab='Time(Hr)', ylab='OD600', main='Growth on Glycolate', bty='n', ylim=c(0, 0.2),cex=1.5, col='gray80')
OD(rpomGlycolate, rpomColor)
OD(vibGlycolate, vibColor)
OD(mariGlycolate, mariColor)

#dev.off()

detach(glycolate_mean)

```


```{r}
ArgGln <- fread('Data/cbs_growth_arginine_gln_malonate.csv') %>% 
    pivot_longer(., cols=2:19, names_to='Organism', values_to='OD') %>% 
    separate(., col='Organism', into=c('Organism', 'Condition', 'Rep')) %>% 
    group_by(Time, Organism, Condition) %>% 
    summarize(mean_OD=mean(OD), sem_OD=sd(OD)/sqrt(length(OD)))
attach(ArgGln)

rpomArg <- Organism=='Rp' & Condition=='arginine'
mariArg <- Organism=='Marivivens' & Condition=='arginine'
vibArg <- Organism=='Vibrio' & Condition=='arginine'
#pdf('ArginineGrowth.pdf')
plot(Time[rpomArg], mean_OD[rpomArg], type='n',  las=1, xlab='Time(h)', ylab='OD600', main='Growth on Arginine', bty='n', ylim=c(0, 0.2),cex=1.5, col='gray80')
OD(rpomArg, rpomColor)
OD(vibArg, vibColor)
OD(mariArg, mariColor)
#dev.off()

rpomGln <- Organism=='Rp' & Condition=='glutamine'
mariGln <- Organism=='Marivivens' & Condition=='glutamine'
vibGln <- Organism=='Vibrio' & Condition=='glutamine'
#pdf('GlutamineGrowth.pdf')
plot(Time[rpomGln], mean_OD[rpomGln], type='n',  las=1, xlab='Time(h)', ylab='OD600', main='Growth on Glutamine', bty='n', ylim=c(0, 0.2),cex=1.5, col='gray80')
OD(rpomGln, rpomColor)
OD(vibGln, vibColor)
OD(mariGln, mariColor)
#dev.off()


detach()
```

```{r}
GluRibose <- fread('Data/cbs_glutamate_ribose.csv')%>%  melt(., id=('Time'), variable.name='Well', value.name='OD600')

GluRibose_layout <- fread('Data/cbs_glutamate_ribose_plateLayout.csv')


GluRibose_growthData <- inner_join(GluRibose, GluRibose_layout , by='Well')


GluRibose_mean <- meanCells(GluRibose_growthData)


GluRibose_mean$mean_OD[GluRibose_mean$Organism!='Blank' & GluRibose_mean$Condition=='Glutamate'] <- GluRibose_mean$mean_OD[GluRibose_mean$Organism!='Blank' & GluRibose_mean$Condition=='Glutamate'] - GluRibose_mean$mean_OD[GluRibose_mean$Organism=='Blank' & GluRibose_mean$Condition=='Glutamate']


attach(GluRibose_mean)

rpomGlu <- Organism=='Rpom' & Condition=='Glutamate'
vibGlu <- Organism=='Vib' & Condition=='Glutamate'
mariGlu <- Organism=='Mari' & Condition=='Glutamate'

#pdf('GlutamateGrowth.pdf')
plot(Time[rpomGlu], mean_OD[rpomGlu], type='n',  las=1, xlab='Time(h)', ylab='OD600', main='Growth on Glutamate', bty='n', ylim=c(0, 0.3),cex=1.5, col='gray80')
OD(rpomGlu, rpomColor)
OD(vibGlu, vibColor)
OD(mariGlu, mariColor)
#dev.off()


detach()
```


```{r}
DMSP <- fread('Data/DSS3,HF1,HF70_DMSP_48-well.csv') %>% 
    pivot_longer(., cols=2:25, names_to='Organism', values_to='OD') %>% 
  na.omit() %>% 
    separate(., col='Organism', into=c('Organism', 'Condition', 'Rep')) %>% 
    group_by(Time, Organism, Condition) %>% 
    summarize(mean_OD=mean(OD), sem_OD=sd(OD)/sqrt(length(OD)))
attach(DMSP)


rpomDMSP <- Organism=='DSS3' & Condition=='DMSP'
mariDMSP <- Organism=='HF1' & Condition=='DMSP'
vibDMSP <- Organism=='HF70' & Condition=='DMSP'
#pdf('DMSP.pdf')
plot(Time[rpomDMSP], mean_OD[rpomDMSP], type='n',  las=1, xlab='Time(h)', ylab='OD600', main='DMSP', bty='n', ylim=c(0, 0.2),xlim=c(0, 20), cex=1.5, col='gray80')
OD(rpomDMSP, rpomColor)
OD(vibDMSP, vibColor)
OD(mariDMSP, mariColor)
#dev.off()
detach()

```


